A method for regularization of evolutionary polynomial regressions

### 2015-08-11 texts/submission_03_asc/eprr.tex
Versão pré-final

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amssymb}
\usepackage{latexsym}
%
%	ALGORITHMS PACKAGES
%
\usepackage{algpseudocode}
\usepackage{algorithm}
%
%	HYPERREFERENCES
%
\usepackage{url}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%	DOCUMENT COMMANDS
%
\newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
\newcommand{\at}[1]{\ensuremath{\!\left(#1\right)}}
\newcommand{\note}[1]{\textbf{\small{[Note: {#1}]}}}
\newcommand{\nth}{\ensuremath{^{th}}}
\newcommand{\review}[1]{\textcolor{blue}{~---~#1}}
\newcommand{\revised}[2]{\review{#1}~\textcolor{red}{#2}}
%
\usepackage{lineno}
%
%	ACRONYMS
%
\usepackage{acronym}
%
%
%
\def\TM{$^{\rm TM}$}

@@

\begin{figure*}[tb]\begin{center}
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{Fig1a.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig1b.pdf}
\\
\includegraphics[width=0.48\textwidth]{Fig1c.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig1d.pdf}
\end{tabular}

@@

\begin{table}[t]
	\footnotesize
	\renewcommand{\arraystretch}{0.55}
	\begin{tabular}{ll|ccc}
		&&& \textbf{error} \\
		\textbf{dataset} & \textbf{method} & \textbf{quantile 25\%} & \textbf{mean} & \textbf{quantile 75\%} \\
		\hline
		Abalone	 & EPRR $\lambda = 0.7$ 	& 0.6392 & 0.6555 & 0.6677 \\
					 & EPRR $\lambda = 0.8$ 	& 0.6408 & 0.6543 & 0.6636 \\
					 & EPRR $\lambda = 0.9$ 	& 0.6481 & 0.6581 & 0.6707 \\
					 & EPRR $\lambda = 1.0$ 	& 0.6542 & 0.6715 & 0.6816 \\
					 & Linear Regression 	& 0.6803 & 0.6927 & 0.7078 \\
					 & SVM (linear kernel) 	& 0.6916 & 0.7044 & 0.7205 \\
					 & Regression Trees 	& 0.7423 & 0.7520 & 0.7621 \\
					 & Random Forest 	& 0.6585 & 0.6695 & 0.6814 \\
					 & Cond. Inference Trees 	& 0.7031 & 0.7126 & 0.7264 \\
		\hline
		Auto-Mpg	 & EPRR $\lambda = 0.7$ 	& 0.3635 & 0.3916 & 0.4147 \\
					 & EPRR $\lambda = 0.8$ 	& 0.3629 & 0.3956 & 0.4228 \\
					 & EPRR $\lambda = 0.9$ 	& 0.3646 & 0.4130 & 0.4215 \\
					 & EPRR $\lambda = 1.0$ 	& 0.3691 & 0.3999 & 0.4057 \\
					 & Linear Regression 	& 0.4071 & 0.4284 & 0.4473 \\
					 & SVM (linear kernel) 	& 0.4116 & 0.4358 & 0.4613 \\
					 & Regression Trees 	& 0.4216 & 0.4501 & 0.4785 \\
					 & Random Forest 	& 0.3318 & 0.3624 & 0.3892 \\
					 & Cond. Inference Trees 	& 0.4063 & 0.4372 & 0.4663 \\
		\hline
		Housing	 & EPRR $\lambda = 0.7$ 	& 0.4412 & 0.6650 & 0.5739 \\
					 & EPRR $\lambda = 0.8$ 	& 0.4241 & 0.5274 & 0.6016 \\
					 & EPRR $\lambda = 0.9$ 	& 0.4354 & 0.5717 & 0.6462 \\
					 & EPRR $\lambda = 1.0$ 	& 0.4477 & 0.5417 & 0.5995 \\
					 & Linear Regression 	& 0.4898 & 0.5313 & 0.5649 \\
					 & SVM (linear kernel) 	& 0.4831 & 0.5469 & 0.6017 \\
					 & Regression Trees 	& 0.4845 & 0.5232 & 0.5720 \\
					 & Random Forest 	& 0.3283 & 0.3679 & 0.4023 \\
					 & Cond. Inference Trees 	& 0.4676 & 0.5080 & 0.5413 \\
		\hline
		Kinematics	 & EPRR $\lambda = 0.7$ 	& 0.6600 & 0.6660 & 0.6720 \\
					 & EPRR $\lambda = 0.8$ 	& 0.6617 & 0.6694 & 0.6751 \\
					 & EPRR $\lambda = 0.9$ 	& 0.6636 & 0.6714 & 0.6761 \\
					 & EPRR $\lambda = 1.0$ 	& 0.7568 & 0.7558 & 0.7739 \\
					 & Linear Regression 	& 0.7672 & 0.7759 & 0.7849 \\
					 & SVM (linear kernel) 	& 0.8074 & 0.8136 & 0.8247 \\
					 & Regression Trees 	& 0.5673 & 0.6021 & 0.5803 \\
					 & Random Forest 	& 0.7386 & 0.7344 & 0.7645 \\
					 & Cond. Inference Trees 	& 0.7558 & 0.7620 & 0.7686
	\end{tabular}
	\caption{Tabular summary results for different regression methods on common datasets. Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.}
\end{table}
Here is described the experiment setup used to gather and summarize the empirical evidence that supports this comparative study of \ac{EPR} and \ac{EPRR}. Evaluation is focused in \ac{rmse} distribution and, besides \ac{EPR} and \ac{EPRR}, also uses several common regression methods and datasets easily accessible in \texttt{R}, the free software environment for statistical computing and graphics~\citep{R-Core-Team:2013aa}\footnote{The datasets and \texttt{R} code used to produce the results and plots in this paper are available online at \url{https://github.com/jpneto/GenAlgPoly}.}. A small consideration on the convergence speed concludes this section.

@@

The \ac{EPRR} method is ranked against several well-known learning algorithms for regression, namely: non-regularized \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa} with linear kernel, Regression Trees~\citep{Therneau:2013aa}, Random Forest~\citep{Strobl:2008aa,Strobl:2007aa} and Conditional Inference Trees~\citep{Hothorn:2006aa}.
%
%	INCLUDE ?
%
%To achieve better error results the \ac{SVM} and Regression Tree parameters are tuned in each dataset.

@@

%
\begin{figure*}[tb]\begin{center}
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{Fig3a.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3b.pdf}
\\
\includegraphics[width=0.48\textwidth]{Fig3c.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3d.pdf}
\end{tabular}
%
\caption{Graphical summary results for different regression methods on common datasets. Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.  The regression methods depicted in these figures are:
1. \ac{EPRR}, $\lambda = 0.7$;
2. \ac{EPRR}, $\lambda = 0.8$;
3. \ac{EPRR}, $\lambda = 0.9$;
4. \ac{EPR} (i.e. $\lambda = 1.0$);
5. Linear Regression;
6. \ac{SVM} (linear kernel);
7. Regression Trees;
8. Random Forest (with 100 trees);
9. Conditional Inference Trees.}
\label{fig:four.datasets.summary}
\label{Housing_dataset_lambda0.8_25runs}
\label{Abalone_dataset_lambda0.8_25runs}
\label{Auto-Mpg_dataset_lambda0.8_25runs}
\label{Kinematics300_lambda0.8_25runs}
\end{center}\end{figure*}

\subsection{Convergence speed}
%
\begin{figure*}[tb]
	\begin{center}
		\includegraphics[width=0.98\textwidth]{Fig4xx.pdf}
		\caption{Learning curve: Error progress for the \textsc{Housing} dataset during a single execution of the genetic algorithm. The figure shows the fitness evolution for different regularization values. The population of each run consists of $200$ polynomials.}
		\label{Abalone_fitnessProgress}
	\end{center}
\end{figure*}
%
Since this work is oriented to the \ac{EPRR} model \ac{rmse} it is necessary to assess how this depends on the number of generations of the \ac{GA}. As illustrated in Figure~\ref{Abalone_fitnessProgress}, the \ac{rmse} quickly drops during the initial $100$ generations. Then, it proceeds slower achieving better solutions only with marginal error reduction. The \ac{EPR} takes more than $300$ generations to achieve similar error results.

@@

Of the regression methods considered \ac{RF} achieves the best results in two out of four datasets. If we compare against non-ensemble methods, \ac{EPR} and \ac{EPRR} outperform the other estimators, with the exception of Regression Trees in the Kinematics dataset.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amssymb}
\usepackage{latexsym}
%
%	ALGORITHMS PACKAGES
%
\usepackage{algpseudocode}
\usepackage{algorithm}
%
%	HYPERREFERENCES
%
\usepackage{url}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%	DOCUMENT COMMANDS
%
\newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
\newcommand{\at}[1]{\ensuremath{\!\left(#1\right)}}
\newcommand{\note}[1]{\textbf{\small{[Note: {#1}]}}}
\newcommand{\nth}{\ensuremath{^{th}}}
\newcommand{\review}[1]{\textcolor{blue}{~---~#1}}
%\newcommand{\revised}[2]{\review{#1}~\textcolor{red}{#2}}
\newcommand{\revised}[2]{}
%
\usepackage{lineno}
%
%	ACRONYMS
%
\usepackage{acronym}
%
%
%
\def\TM{$^{\rm TM}$}

@@

\begin{figure*}[tb]\begin{center}
\begin{tabular}{cc}
\includegraphics[width=0.475\textwidth]{Fig1a.pdf}
&
\includegraphics[width=0.475\textwidth]{Fig1b.pdf}
\\
\includegraphics[width=0.475\textwidth]{Fig1c.pdf}
&
\includegraphics[width=0.475\textwidth]{Fig1d.pdf}
\end{tabular}

@@

\begin{table}[t]
	\footnotesize
	\renewcommand{\arraystretch}{0.55}
	\begin{tabular}{ll|ccc}
		&&& \textbf{error} \\
		\textbf{dataset} & \textbf{method} & \textbf{quantile 25\%} & \textbf{mean} & \textbf{quantile 75\%} \\
		\hline
		Abalone	 & EPRR $\lambda = 0.7$ 	& 0.6392 & 0.6555 & 0.6677 \\
					 & EPRR $\lambda = 0.8$ 	& 0.6408 & 0.6543 & 0.6636 \\
					 & EPRR $\lambda = 0.9$ 	& 0.6481 & 0.6581 & 0.6707 \\
					 & EPRR $\lambda = 1.0$ 	& 0.6542 & 0.6715 & 0.6816 \\
					 & Linear Regression 	& 0.6803 & 0.6927 & 0.7078 \\
					 & SVM (linear kernel) 	& 0.6916 & 0.7044 & 0.7205 \\
					 & Regression Trees 	& 0.7423 & 0.7520 & 0.7621 \\
%					 & Random Forest 	& 0.6585 & 0.6695 & 0.6814 \\
					 & Cond. Inference Trees 	& 0.7031 & 0.7126 & 0.7264 \\
		\hline
		Auto-Mpg	 & EPRR $\lambda = 0.7$ 	& 0.3635 & 0.3916 & 0.4147 \\
					 & EPRR $\lambda = 0.8$ 	& 0.3629 & 0.3956 & 0.4228 \\
					 & EPRR $\lambda = 0.9$ 	& 0.3646 & 0.4130 & 0.4215 \\
					 & EPRR $\lambda = 1.0$ 	& 0.3691 & 0.3999 & 0.4057 \\
					 & Linear Regression 	& 0.4071 & 0.4284 & 0.4473 \\
					 & SVM (linear kernel) 	& 0.4116 & 0.4358 & 0.4613 \\
					 & Regression Trees 	& 0.4216 & 0.4501 & 0.4785 \\
%					 & Random Forest 	& 0.3318 & 0.3624 & 0.3892 \\
					 & Cond. Inference Trees 	& 0.4063 & 0.4372 & 0.4663 \\
		\hline
		Housing	 & EPRR $\lambda = 0.7$ 	& 0.4412 & 0.6650 & 0.5739 \\
					 & EPRR $\lambda = 0.8$ 	& 0.4241 & 0.5274 & 0.6016 \\
					 & EPRR $\lambda = 0.9$ 	& 0.4354 & 0.5717 & 0.6462 \\
					 & EPRR $\lambda = 1.0$ 	& 0.4477 & 0.5417 & 0.5995 \\
					 & Linear Regression 	& 0.4898 & 0.5313 & 0.5649 \\
					 & SVM (linear kernel) 	& 0.4831 & 0.5469 & 0.6017 \\
					 & Regression Trees 	& 0.4845 & 0.5232 & 0.5720 \\
%					 & Random Forest 	& 0.3283 & 0.3679 & 0.4023 \\
					 & Cond. Inference Trees 	& 0.4676 & 0.5080 & 0.5413 \\
		\hline
		Kinematics	 & EPRR $\lambda = 0.7$ 	& 0.6600 & 0.6660 & 0.6720 \\
					 & EPRR $\lambda = 0.8$ 	& 0.6617 & 0.6694 & 0.6751 \\
					 & EPRR $\lambda = 0.9$ 	& 0.6636 & 0.6714 & 0.6761 \\
					 & EPRR $\lambda = 1.0$ 	& 0.7568 & 0.7558 & 0.7739 \\
					 & Linear Regression 	& 0.7672 & 0.7759 & 0.7849 \\
					 & SVM (linear kernel) 	& 0.8074 & 0.8136 & 0.8247 \\
					 & Regression Trees 	& 0.5673 & 0.6021 & 0.5803 \\
%					 & Random Forest 	& 0.7386 & 0.7344 & 0.7645 \\
					 & Cond. Inference Trees 	& 0.7558 & 0.7620 & 0.7686
	\end{tabular}
	\caption{Tabular summary results for different regression methods on common datasets.}
\end{table}
Here is described the experiment setup used to gather and summarize the empirical evidence that supports this comparative study of \ac{EPR} and \ac{EPRR}. Evaluation is focused in \ac{rmse} distribution and, besides \ac{EPR} and \ac{EPRR}, also uses several common regression methods and datasets easily accessible in \texttt{R}, the free software environment for statistical computing and graphics~\citep{R-Core-Team:2013aa}. Datasets and code used to produce the results and plots in this paper are available at \url{https://github.com/jpneto/GenAlgPoly}. A small consideration on the convergence speed concludes this section.

@@

The \ac{EPRR} method is ranked against several well-known learning algorithms for regression, namely: non-regularized \ac{EPR}, Linear Regression, Support Vector Machines~\citep{Meyer:2012aa} with linear kernel, Regression Trees~\citep{Therneau:2013aa} and Conditional Inference Trees~\citep{Hothorn:2006aa,Strobl:2008aa,Strobl:2007aa}. No ensemble regression method was used in this comparison since these methods average between several models while \ac{EPRR} and the selected methods produce a single model.
%
%	INCLUDE ?
%
%To achieve better error results the \ac{SVM} and Regression Tree parameters are tuned in each dataset.

@@

%
\begin{figure*}[tb]\begin{center}
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{Fig3a.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3b.pdf}
\\
\includegraphics[width=0.48\textwidth]{Fig3c.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3d.pdf}
\end{tabular}
%
\caption{Graphical summary results for different regression methods on common datasets. 
The regression methods depicted in these figures are:
1. \ac{EPRR}, $\lambda = 0.7$;
2. \ac{EPRR}, $\lambda = 0.8$;
3. \ac{EPRR}, $\lambda = 0.9$;
4. \ac{EPR} (i.e. $\lambda = 1.0$);
5. Linear Regression;
6. \ac{SVM} (linear kernel);
7. Regression Trees;
%8. Random Forest (with 100 trees);
8. Conditional Inference Trees.}
\label{fig:four.datasets.summary}
\label{Housing_dataset_lambda0.8_25runs}
\label{Abalone_dataset_lambda0.8_25runs}
\label{Auto-Mpg_dataset_lambda0.8_25runs}
\label{Kinematics300_lambda0.8_25runs}
\end{center}\end{figure*}

\subsection{Convergence speed}
%
\begin{figure*}[tb]
	\begin{center}
		\includegraphics[width=0.98\textwidth]{Fig4.pdf}
		\caption{Learning curve: Error progress for the \textsc{Housing} dataset during a single execution of the genetic algorithm. The figure shows the fitness evolution for different regularization values. The population of each run consists of $200$ polynomials.}
		\label{Abalone_fitnessProgress}
	\end{center}
\end{figure*}
%
Since this work is oriented to the \ac{EPRR} model \ac{rmse} it is necessary to assess how this depends on the number of generations of the \ac{GA}. As illustrated in Figure~\ref{Abalone_fitnessProgress}, the \ac{rmse} quickly drops during the initial $100$ generations. Then, it proceeds slower achieving better solutions only with marginal error reduction. The \ac{EPR} takes more than $300$ generations to achieve similar error results.

@@

Of the regression methods considered \ac{EPRR} achieves the best results in three out of four datasets, the exception being Regression Trees in the Kinematics dataset.
